---
title: "Clean and Summarize Data"
authors: "Savannah Weaver, Corey Callaghan"
date: "01/01/2020"
output: 
  github_document:
    toc: true
    toc_depth: 1
---
# Load Packages

```{r setup, echo=T, results='hide', message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if (!require("readr")) install.packages("readr")
library("readr")
if (!require("dplyr")) install.packages("dplyr")
library("dplyr")
if (!require("tidyr")) install.packages("tidyr")
library("tidyr")
if (!require("ggplot2")) install.packages("ggplot2")
library("ggplot2")
if (!require("sf")) install.packages("sf")
library("sf")
if (!require("lubridate")) install.packages("lubridate")
library("lubridate")
if (!require("lwgeom")) install.packages("lwgeom")
library("lwgeom")
if (!require("rnaturalearth")) install.packages("rnaturalearth")
library("rnaturalearth")
if (!require("rnaturalearthdata")) install.packages("rnaturalearthdata")
library("rnaturalearthdata")
if (!require("patchwork")) install.packages("patchwork")
library("patchwork")
if (!require("mgcv")) install.packages("mgcv")
library("mgcv")
if (!require("rgeos")) install.packages("rgeos")
library("rgeos")
if (!require("broom")) install.packages("broom")
library("broom")
library(extrafont)
loadfonts()
```


# Reading In Data

```{r}
all.species <- readRDS("data_for_analysis/call_dat.RDS")

# find freq max and min across species and individuals based on single calls
max(all.species$peak_freq_hz) #5062.5
min(all.species$peak_freq_hz) #344.5
```

Select pertinent variables only:

```{r}
data <- all.species %>%
  dplyr::select(id, species,
                lat, lng,
                delta_time_s,
                peak_freq_hz,
                capture_time_in_zone) %>%
  mutate(date = as.Date(capture_time_in_zone))
```

Load expected number of observations for each species:

```{r}
expected <- read.table(textConnection("
Species                   Unique_IDs
Crinia_insignifera        123
Crinia_parinsignifera     221
Limnodynastes_dorsalis    82
Limnodynastes_peronii     232
Litoria_chloris           48
Litoria_xanthomera        56
"), header=TRUE)
```

Check expected values against actual:

```{r}
test <- data %>%
  group_by(species) %>%
  summarise(number=length(unique(id)))

compare <- expected$Unique_IDs == test$number
compare
```


# Get Date Ranges

Export the date range of the recordings we analyzed:

```{r}
dates <- data %>%
  group_by(species) %>%
  mutate(earliest = min(date)) %>%
  mutate(latest = max(date)) %>%
  dplyr::select(species, earliest, latest) %>%
  distinct()

#write_csv(dates, "exported_tables/analyzed_sample_dates.csv")
```


# Calculate Summary Data

For each individual/recording, we want to get the following data:
-number of calls/notes analyzed
-mean and standard deviation of call/note duration
-mean and standard deviation of peak frequency

```{r}
per.individual <- data %>%
  group_by(id, species, lat, lng, date) %>%
  add_tally() %>%
  summarise(n_calls = mean(n),
            ind_mean_freq=mean(peak_freq_hz),
            ind_sd_freq=sd(peak_freq_hz),
            ind_mean_time=mean(delta_time_s),
            ind_sd_time=sd(delta_time_s))
```

Also calculate these values for each species overall, export, and add to dataframe:

```{r}
sp_mean_sd <- per.individual %>%
  group_by(species) %>%
  summarize(sp_mean_freq=mean(ind_mean_freq),
            sp_sd_freq=sd(ind_mean_freq),
            sp_mean_time=mean(ind_mean_time),
            sp_sd_time=sd(ind_mean_time))
#write_csv(sp_mean_sd, "data_for_analysis/species_data.csv")

per.ind.sp <- left_join(per.individual,
                        sp_mean_sd,
                        by = "species")

# check proper joining, each should =6
length(unique(per.ind.sp$sp_mean_freq))
length(unique(per.ind.sp$sp_mean_time))
length(unique(per.ind.sp$sp_sd_freq))
length(unique(per.ind.sp$sp_sd_freq))

# min and max frequencies based on individual means
max(per.ind.sp$ind_mean_freq) #5015.633
min(per.ind.sp$ind_mean_freq) #430.7
```


# Visualize Data Distribution

Make histograms for call duration and peak frequency:

```{r}
# mean_freq
ggplot(per.ind.sp, aes(x=ind_mean_freq)) +
  geom_histogram(color="black", fill="steelblue", bins=50) +
  theme_bw() +
  facet_wrap(~species, scales="free") +
  xlab("Mean Frequency")

# mean_time
ggplot(per.ind.sp, aes(x=ind_mean_time))+
  geom_histogram(color="black", fill="steelblue", bins=50)+
  theme_bw()+
  facet_wrap(~species, scales="free")+
  xlab("Mean Time")
```

Use one plot to compare call variables and their distribution among species:

```{r}
per.ind.sp %>%
  rename(Frequency = ind_mean_freq) %>%
  rename(Time = ind_mean_time) %>%
  ggplot(., aes(x = Time,
                y = Frequency/1000,
                color = as.factor(species),
                shape = species)) +
  geom_point(size = 2) + 
  scale_colour_manual(name = "Species",
                      labels = c(expression(italic("Crinia insignifera")),
                                 expression(italic("Crinia parinsignifera")),
                                 expression(italic("Limnodynastes dorsalis")),
                                 expression(italic("Limnodynastes peronii")),
                                 expression(italic("Litoria chloris")),
                                 expression(italic("Litoria xanthomera"))),
                      values = c(rep(c(rgb(168/255,221/255,181/255),
                                       rgb(67/255,162/255,202/255)),3))) +
  scale_shape_manual(name = "Species",
                     labels = c(expression(italic("Crinia insignifera")),
                                 expression(italic("Crinia parinsignifera")),
                                 expression(italic("Limnodynastes dorsalis")),
                                 expression(italic("Limnodynastes peronii")),
                                 expression(italic("Litoria chloris")),
                                 expression(italic("Litoria xanthomera"))),
                     values = c(15,15,16,16,17,17)) +
  stat_ellipse(size = 0.5,
               show.legend = FALSE) +
  theme_classic() +
  xlab("Mean Individual Call Duration (seconds)") +
  ylab("Mean Individual Peak Frequency (kHz)") +
  theme(text = element_text(color = "black", 
                            family = "sans",
                            size = 12),
        axis.text = element_text(color = "black", 
                            family = "sans",
                            size = 10),
        legend.position = c(0.8, 0.8), 
        legend.text.align = 0,
        legend.title = element_blank()) -> compare_species_fig

compare_species_fig

#ggsave(filename = "compare_species_fig.png",
 #      plot = compare_species_fig,
  #     path = "./exported_figures",
   #    device = "png",
    #   dpi = 2400,#5.5MB
     #  width = 7, height = 5)
```


# Visualize Species Ranges

Get geographic data:

```{r}
# get geography of Australia
aus <- st_as_sf(ne_countries(country = 'australia', scale = 'medium'))

# get each species range map
load("range_maps/Crinia_insignifera.RData")
crinia_insignifera <- st_as_sf(range_map) 

load("range_maps/Crinia_parinsignifera.RData")
crinia_parinsignifera <- st_as_sf(range_map) 

load("range_maps/Litoria_chloris.RData")
litoria_chloris <- st_as_sf(range_map) 

load("range_maps/Litoria_xanthomera.RData")
litoria_xanthomera <- st_as_sf(range_map) 

load("range_maps/Limnodynastes_peronii.RData")
limnodynastes_peronii <- st_as_sf(range_map) 

load("range_maps/Limnodynastes_dorsalis.RData")
limnodynastes_dorsalis <- st_as_sf(range_map) 
```

Plot all species ranges on Australian continent map:

```{r}
all_ranges_plot <- ggplot() +
  theme_classic() +
  geom_sf(data = aus) +
  # limnodynastes dorsalis
  geom_sf(data = limnodynastes_dorsalis, 
          fill = "tan1", color = "tan1", alpha = 0.6) +
  # limnodynastes peronii
  geom_sf(data = limnodynastes_peronii, 
          fill = "indianred1", color = "indianred1", alpha = 0.6) + 
  # crinia parinsignifera
  geom_sf(data = crinia_parinsignifera, 
          fill = "steelblue1", color = "steelblue1", alpha = 0.6) +
  # crinia insignifera
  geom_sf(data = crinia_insignifera, 
          fill = "darkseagreen", color = "darkseagreen", alpha = 0.7) +
  # litoria chloris
  geom_sf(data = litoria_chloris, 
          fill = "gray", color = "gray", alpha = 0.7) +
  # litoria xanthomera
  geom_sf(data = litoria_xanthomera, 
          fill = "black", color = "black", alpha = 0.5) +
  xlim(110,155) + ylim(45, 10) +
  ggtitle("All Species Ranges")
all_ranges_plot
```

For Figure 1, map each genus-pair of species:

```{r}
crinia_ranges_plot <- ggplot() +
  theme_classic() +
  geom_sf(data = aus) +
  # crinia parinsignifera
  geom_sf(data = crinia_parinsignifera, 
          fill = rgb(67/255,162/255,202/255), alpha = 0.8, 
          color = rgb(67/255,162/255,202/255)) +
  # crinia insignifera
  geom_sf(data = crinia_insignifera, 
          fill = rgb(168/255,221/255,181/255), alpha = 0.8, 
          color = rgb(168/255,221/255,181/255)) +
  xlim(110,155) + ylim(45, 10) +
  ggtitle("Crinia Species Ranges")

crinia_ranges_plot
#ggsave(filename = "crinia_ranges_plot.png",
 #      plot = crinia_ranges_plot,
  #     path = "exported_figures",
   #    device = "png",
    #   dpi = 1200)

litoria_ranges_plot <- ggplot() +
  theme_classic() +
  geom_sf(data = aus) +
  # litoria chloris
  geom_sf(data = litoria_chloris, 
          fill = rgb(168/255,221/255,181/255), alpha = 0.8, 
          color = rgb(168/255,221/255,181/255)) +
  # litoria xanthomera
  geom_sf(data = litoria_xanthomera, 
          fill = rgb(67/255,162/255,202/255), alpha = 0.8, 
          color = rgb(67/255,162/255,202/255)) +
  xlim(110,155) + ylim(45, 10) +
  ggtitle("Litoria Species Ranges")

litoria_ranges_plot
#ggsave(filename = "litoria_ranges_plot.png",
 #      plot = litoria_ranges_plot,
  #     path = "exported_figures",
   #    device = "png",
    #   dpi = 1200)

limnodynastes_ranges_plot <- ggplot() +
  theme_classic() +
  geom_sf(data = aus) +
  # limnodynastes dorsalis
  geom_sf(data = limnodynastes_dorsalis, 
          fill = rgb(168/255,221/255,181/255), alpha = 0.8, 
          color = rgb(168/255,221/255,181/255)) +
  # limnodynastes peronii
  geom_sf(data = limnodynastes_peronii,  
          fill = rgb(67/255,162/255,202/255), alpha = 0.8, 
          color = rgb(67/255,162/255,202/255)) +
  xlim(110,155) + ylim(45, 10) +
  ggtitle("Limnodynastes Species Ranges")

limnodynastes_ranges_plot
#ggsave(filename = "limnodynastes_ranges_plot.png",
 #      plot = limnodynastes_ranges_plot,
  #     path = "exported_figures",
   #    device = "png",
    #   dpi = 1200)
```


# Map Species Ranges With Locations of Recordings Used

These are for supplementary figure 1.

Crinia parinsignifera:

```{r}
cp_records <- per.individual %>%
  dplyr::select(id, species, lat, lng) %>%
  dplyr::filter(species == "Crinia parinsignifera")
cp_records_sf <- st_as_sf(cp_records, 
                          coords = c("lng", "lat"), 
                          crs = 4326, agr = "constant")

cp_rec_plot <- ggplot() +
  theme_classic() +
  geom_sf(data = aus) +
  geom_sf(data = crinia_parinsignifera, 
          fill = rgb(67/255,162/255,202/255), alpha = 0.8, 
          color = rgb(67/255,162/255,202/255)) +
  geom_sf(data = cp_records_sf, 
          size = 2, shape = 19, 
          color = "black") +
  ggtitle(expression(paste(italic("Crinia parinsignifera")))) +
  theme(title = element_text(color = "black", 
                            family = "sans",
                            size = 12),
        axis.text = element_text(color = "black", 
                            family = "sans",
                            size = 10)) +
  scale_x_continuous(breaks = c(140,145,150),
                     limits = c(137, 153)) +
  scale_y_continuous(breaks = c(-25, -30, -35),
                     limits = c(-39, -23))

cp_rec_plot
#ggsave(filename = "cp_rec_plot.jpeg",
 #      plot = cp_rec_plot,
  #     path = "./exported_figures/ind_sp_maps",
   #    device = "jpeg",
    #   width = 3, height = 3,
     #  dpi = 1200)
```

Crinia insignifera:

```{r}
ci_records <- per.individual %>%
  dplyr::select(id, species, lat, lng) %>%
  dplyr::filter(species == "Crinia insignifera")
ci_records_sf <- st_as_sf(ci_records, 
                          coords = c("lng", "lat"), 
                          crs = 4326, agr = "constant")

ci_rec_plot <- ggplot() +
  theme_classic() +
  geom_sf(data = aus) +
  geom_sf(data = crinia_insignifera, 
          fill = rgb(168/255,221/255,181/255), alpha = 0.8, 
          color = rgb(168/255,221/255,181/255)) +
  geom_sf(data = ci_records_sf, 
          size = 2, shape = 19, 
          color = "black") +
  ggtitle(expression(paste(italic("Crinia insignifera")))) +
  theme(title = element_text(color = "black", 
                            family = "sans",
                            size = 12),
        axis.text = element_text(color = "black", 
                            family = "sans",
                            size = 10)) +
  scale_x_continuous(breaks = c(114,116),
                     limits = c(113, 117)) +
  scale_y_continuous(breaks = c(-30, -32, -34),
                     limits = c(-34, -30))
ci_rec_plot
#ggsave(filename = "ci_rec_plot.jpeg",
 #      plot = ci_rec_plot,
  #     path = "./exported_figures/ind_sp_maps",
   #    device = "jpeg",
    #   width = 3, height = 3,
     #  dpi = 1200)
```

Litoria chloris:

```{r}
lc_records <- per.individual %>%
  dplyr::select(id, species, lat, lng) %>%
  dplyr::filter(species == "Litoria chloris")
lc_records_sf <- st_as_sf(lc_records, 
                          coords = c("lng", "lat"), 
                          crs = 4326, agr = "constant")

lc_rec_plot <- ggplot() +
  theme_classic() +
  geom_sf(data = aus) +
  geom_sf(data = litoria_chloris, 
          fill = rgb(168/255,221/255,181/255), alpha = 0.8, 
          color = rgb(168/255,221/255,181/255)) +
  geom_sf(data = lc_records_sf, 
          size = 2, shape = 19, 
          color = "black") +
  ggtitle(expression(paste(italic("Litoria chloris")))) +
  theme(title = element_text(color = "black", 
                            family = "sans",
                            size = 12),
        axis.text = element_text(color = "black", 
                            family = "sans",
                            size = 10)) +
  scale_x_continuous(breaks = c(145,150,155),
                     limits = c(143, 159)) +
  scale_y_continuous(breaks = c(-20, -25, -30),
                     limits = c(-35, -19))
lc_rec_plot
#ggsave(filename = "lc_rec_plot.jpeg",
 #      plot = lc_rec_plot,
  #     path = "./exported_figures/ind_sp_maps",
   #    device = "jpeg",
    #   width = 3, height = 3,
     #  dpi = 1200)
```

Litoria xanthomera:

```{r}
lx_records <- per.individual %>%
  dplyr::select(id, species, lat, lng) %>%
  dplyr::filter(species == "Litoria xanthomera")
lx_records_sf <- st_as_sf(lx_records, 
                          coords = c("lng", "lat"), 
                          crs = 4326, agr = "constant")

lx_rec_plot <- ggplot() +
  theme_classic() +
  geom_sf(data = aus) +
  geom_sf(data = litoria_xanthomera, 
          fill = rgb(67/255,162/255,202/255), alpha = 0.8, 
          color = rgb(67/255,162/255,202/255)) +
  geom_sf(data = lx_records_sf, 
          size = 2, shape = 19, 
          color = "black") +
  ggtitle(expression(paste(italic("Litoria xanthomera")))) +
  theme(title = element_text(color = "black", 
                            family = "sans",
                            size = 12),
        axis.text = element_text(color = "black", 
                            family = "sans",
                            size = 10)) +
  scale_x_continuous(breaks = c(144,146,148),
                     limits = c(144, 148)) +
  scale_y_continuous(breaks = c(-16, -18),
                     limits = c(-19.5, -15.5))
lx_rec_plot
#ggsave(filename = "lx_rec_plot.jpeg",
 #      plot = lx_rec_plot,
  #     path = "./exported_figures/ind_sp_maps",
   #    device = "jpeg",
    #   width = 3, height = 3,
     #  dpi = 1200)
```

Limnodynastes dorsalis:

```{r}
ld_records <- per.individual %>%
  dplyr::select(id, species, lat, lng) %>%
  dplyr::filter(species == "Limnodynastes dorsalis")
ld_records_sf <- st_as_sf(ld_records, 
                          coords = c("lng", "lat"), 
                          crs = 4326, agr = "constant")

ld_rec_plot <- ggplot() +
  theme_classic() +
  geom_sf(data = aus) +
 geom_sf(data = limnodynastes_dorsalis, 
          fill = rgb(168/255,221/255,181/255), alpha = 0.8, 
          color = rgb(168/255,221/255,181/255)) +
  geom_sf(data = ld_records_sf, 
          size = 2, shape = 19, 
          color = "black") +
  ggtitle(expression(paste(italic("Limnodynastes dorsalis")))) +
  theme(title = element_text(color = "black", 
                            family = "sans",
                            size = 12),
        axis.text = element_text(color = "black", 
                            family = "sans",
                            size = 10)) +
  scale_x_continuous(breaks = c(114,118,122),
                     limits = c(113, 125)) +
  scale_y_continuous(breaks = c(-28, -32, -36),
                     limits = c(-36, -27))
ld_rec_plot
#ggsave(filename = "ld_rec_plot.jpeg",
 #      plot = ld_rec_plot,
  #     path = "./exported_figures/ind_sp_maps",
   #    device = "jpeg",
    #   width = 3, height = 3,
     #  dpi = 1200)
```

Limnodynastes peronii:

```{r}
lp_records <- per.individual %>%
  dplyr::select(id, species, lat, lng) %>%
  dplyr::filter(species == "Limnodynastes peronii")
lp_records_sf <- st_as_sf(lp_records, 
                          coords = c("lng", "lat"), 
                          crs = 4326, agr = "constant")

lp_rec_plot <- ggplot() +
  theme_classic() +
  geom_sf(data = aus) +
  geom_sf(data = limnodynastes_peronii,  
          fill = rgb(67/255,162/255,202/255), alpha = 0.8, 
          color = rgb(67/255,162/255,202/255)) +
  geom_sf(data = lp_records_sf, 
          size = 2, shape = 19, 
          color = "black") +
  ggtitle(expression(paste(italic("Limnodynastes peronii")))) +
  theme(title = element_text(color = "black", 
                            family = "sans",
                            size = 12),
        axis.text = element_text(color = "black", 
                            family = "sans",
                            size = 10)) +
  scale_x_continuous(breaks = c(135,145,155),
                     limits = c(130, 160)) +
  scale_y_continuous(breaks = c(-20, -30, -40),
                     limits = c(-43, -13))
lp_rec_plot
#ggsave(filename = "lp_rec_plot.jpeg",
 #      plot = lp_rec_plot,
  #     path = "./exported_figures/ind_sp_maps",
   #    device = "jpeg",
    #   width = 3, height = 3,
     #  dpi = 1200)
```




# Temperature Model

Scale data (z-score) to compare between species:

```{r}
scaled_dat <- per.ind.sp %>%
  group_by(species) %>%
  mutate(z_freq=((ind_mean_freq-sp_mean_freq)/sp_sd_freq)) %>%
  mutate(z_time=((ind_mean_time-sp_mean_time)/sp_sd_time))
```

Load in temperature data and join:

```{r}
temp_dat <- readRDS("data_for_analysis/recording_temp_estimates.RDS")

data_with_temp <- left_join(scaled_dat,
                            temp_dat,
                            by = c("id", "species"))
```

Plot call variables ~ temperature:

```{r}
data_with_temp %>%
  rename(Frequency = z_freq) %>%
  rename('Call Duration' = z_time) %>%
  gather(key = "Variable", value = "value", c(Frequency, 'Call Duration')) %>%
  ggplot(., aes(x = Hourly_temp, y = value, color = species))+
  geom_smooth(method = "lm", formula = y ~ x, fill = NA)+
  theme_classic()+
  scale_colour_manual(name = "Species",
                        values = c("tan1","indianred1",
                                   "steelblue1","darkseagreen",
                                   "gray","black")) +
  theme(axis.text = element_text(color = "black"))+
  facet_wrap(~Variable, scale = "free")+
  ylab("Z-Score of Call Values")+
  xlab("Hourly Temperature (degrees Celsius)")

data_with_temp %>%
  rename(Frequency = ind_mean_freq) %>%
  rename('Call Duration' = ind_mean_time) %>%
  gather(key = "Variable", value = "value", c(Frequency, 'Call Duration')) %>%
  ggplot(., aes(x = Hourly_temp, y = value, color = species))+
  geom_smooth(method = "lm", formula = y ~ x, fill = NA)+
  theme_classic()+
  scale_colour_manual(name = "Species",
                        values = c("tan1","indianred1",
                                   "steelblue1","darkseagreen",
                                   "gray","black")) +
  theme(axis.text = element_text(color = "black"))+
  facet_wrap(~Variable, scale = "free")+
  ylab("Call Value")+
  xlab("Hourly Temperature (degrees Celsius)")
```

We use the raw call variable values in the models:

```{r}
# call duration
time_by_temp_glm <- glm(data = data_with_temp, 
                      ind_mean_time ~ Hourly_temp)
summary(time_by_temp_glm)

# frequency
freq_by_temp_glm <- glm(data = data_with_temp, 
                      ind_mean_freq ~ Hourly_temp)
summary(freq_by_temp_glm)
```

Export summary data:

```{r}
#write_csv(data_with_temp, "data_for_analysis/summary_data.csv")
```
