---
title: "Function to Get Call Variation Based on Resampling"
authors: "Savannah Weaver, Corey Callaghan"
date: "01/01/2020"
output: 
  github_document:
    toc: true
    toc_depth: 1
---

# Load Packages

```{r setup, echo=T, results='hide', message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if (!require("readr")) install.packages("readr")
library("readr")
if (!require("dplyr")) install.packages("dplyr")
library("dplyr")
if (!require("tidyr")) install.packages("tidyr")
library("tidyr")
if (!require("ggplot2")) install.packages("ggplot2")
library("ggplot2")
if (!require("sf")) install.packages("sf")
library("sf")
if (!require("lubridate")) install.packages("lubridate")
library("lubridate")
if (!require("lwgeom")) install.packages("lwgeom")
library("lwgeom")
```


# Read In Data

```{r, echo=T, message=FALSE}
data <- read_csv("data_for_analysis/summary_data.csv")
```


# Make a Resampling/Bootstrapping Function

For each species, we draw 1000 samples of n observations (n = 3 to the number of individuals sampled for that species). For each sample, we measure the standard deviation of peak frequency and standard deviation of call duration, and note the number of individuals in that sample

The summary_function computes the standard deviations of peak frequency and call duration for a given sample.

```{r}
summary_function <- function(draw, species_x){
  
  # do for each species
  sample_dat <- data %>%
    dplyr::filter(species==species_x) %>%
    ungroup() %>%
    sample_n(draw)
  
  # get summary data
  summary_dat <- data.frame(sd_time=sd(sample_dat$ind_mean_time),
                            sd_freq=sd(sample_dat$ind_mean_freq)) %>%
    mutate(n_samples=draw)
  
  # return data
  return(summary_dat)
}
```

The boot_function repeats summary_function 1000 times for each "draw" of n observations.

```{r}
boot_function <- function(draw, species_x){
  
  # do 1000 samples for each draw and put into one dataframe
  # currently set to 10 for testing
  summary_dat <- do.call("rbind", purrr::rerun(1000, summary_function(draw, species_x)))
  # return dataframe 
  return(summary_dat)
}
```

The species_function applies the above functions to a given species and saves the results:

```{r}
species_function <- function(species_name){
  
  # get the data and do function for one species
  data_to_analyze <- data %>%
    dplyr::filter(species==species_name)
  
  # make a vector of list of possible n values
  vector_of_draws <- rep(3:48, by=1)
  
  # apply function to vector
  list_of_results <- lapply(vector_of_draws, 
                            boot_function, 
                            species_x=species_name)
  
  # put results into dataframe
  df_of_results <- bind_rows(list_of_results) %>%
    mutate(species=species_name)
  
  # return dataframe of results
  return(df_of_results)
}
```

This code chunk runs the above functions for each species and saves the results:

```{r}
# do everything for every species
species_sd_summary_list <- lapply(unique(data$species),
                                  function(x){species_function(x)})

# put all data for each species into one dataframe
species_sd_summary_df <- bind_rows(species_sd_summary_list)

# export
saveRDS(species_sd_summary_df,
        "data_for_analysis/standard_deviation_by_sample_size_data.RDS") 
```

The exported RDS file is analyzed in the "3_resampling_analysis.Rmd" file.