---
title: "Analysis of the Individual Resampling Dataset"
authors: "Savannah Weaver, Corey Callaghan"
date: "01/01/2020"
output: 
  github_document:
    toc: true
    toc_depth: 1
---

# Load Packages

```{r setup, echo=T, results='hide', message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if (!require("readr")) install.packages("readr")
library("readr")
if (!require("dplyr")) install.packages("dplyr")
library("dplyr")
if (!require("tidyr")) install.packages("tidyr")
library("tidyr")
if (!require("ggplot2")) install.packages("ggplot2")
library("ggplot2")
if (!require("lubridate")) install.packages("lubridate")
library("lubridate")
if (!require("lwgeom")) install.packages("lwgeom")
library("lwgeom")
if (!require("lme4")) install.packages("lme4")
library("lme4")
if (!require("lmerTest")) install.packages("lmerTest")
library("lmerTest")
if (!require("broom")) install.packages("broom")
library("broom")
if (!require("mgcv")) install.packages("mgcv")
library("mgcv")
if (!require("nlme")) install.packages("nlme")
library("nlme")
library(extrafont)
loadfonts()
```


# Reading In Data

Get data we made with functions in file 2, and check for correct uploading:

```{r, echo=T, message=FALSE}
data <- read_rds("data_for_analysis/standard_deviation_by_sample_size_data.RDS")
range(data$n_samples)
```

Load in the species summary data and check format:

```{r, echo=T, message=FALSE}
species_values <- read_csv("data_for_analysis/species_data.csv")
species_values
```


# Goal

Investigate how much variation there is within each species call, how many individuals *should* be measured to account for that variation in species definitions, and whether these values differ between species.


# Plot Each Species Separately

This figure shows all data.

```{r}
data %>%
  rename(Frequency = sd_freq) %>%
  rename("Call Duration" = sd_time) %>%
  gather(key = "Variable", value = "value", c(Frequency, "Call Duration")) %>%
  ggplot(., aes(x = n_samples, y = value)) +
  geom_jitter() +
  theme_classic() +
  theme(axis.text = element_text(color = "black")) +
  facet_wrap(species ~ Variable, scale = "free") +
  ylab("Standard Deviation") +
  xlab("Number of Individuals in a Sample")
```


# Plot All Species Together

Scale and join necessary data:

```{r}
# get values
actual <- species_values

# join and scale
scale_all <- data %>%
  left_join(., actual,
            by = "species") %>%
  group_by(species) %>%
  mutate(z_sd_time= ((sd_time) - sp_sd_time) / sd(sd_time)) %>%
  mutate(z_sd_freq= ((sd_freq) - sp_sd_freq) / sd(sd_freq)) %>%
  group_by(species, n_samples) %>%
  mutate(max_z_time = max(abs(z_sd_time))) %>%
  mutate(max_z_freq = max(abs(z_sd_freq)))

# test correct calculations
scale_all %>% 
  group_by(n_samples,species) %>% 
  summarise(test_time = mean(max_z_time),
            test_freq = mean(max_z_freq))

```

This figure only plots the **maximum** standard deviations, relativized to the actual standard deviation for each species we measured.

```{r}
max_points <- scale_all %>%
  group_by(species) %>%
  rename("Peak Frequency" = 'max_z_freq') %>%
  rename("Call Duration" = 'max_z_time') %>%
  gather(key = "Variable", 
         value = "value", 
         c("Peak Frequency", 
           "Call Duration")) %>%
  dplyr::select(species, value, Variable, n_samples) %>%
  distinct() %>% # makes much smaller dataframe to plot (and quicker)
  ggplot() +
  geom_point(aes(x = n_samples, y = value, 
                 color = as.factor(species),
                 shape = as.factor(species)),
             size = 2, alpha = 0.4) +
  scale_colour_manual(labels = c(expression(italic("Crinia insignifera")),
                                 expression(italic("Crinia parinsignifera")),
                                 expression(italic("Limnodynastes dorsalis")),
                                 expression(italic("Limnodynastes peronii")),
                                 expression(italic("Litoria chloris")),
                                 expression(italic("Litoria xanthomera"))),
                      values = c(rep(c(rgb(168/255,221/255,181/255),
                                       rgb(67/255,162/255,202/255)),3))) +
  scale_shape_manual(labels = c(expression(italic("Crinia insignifera")),
                                 expression(italic("Crinia parinsignifera")),
                                 expression(italic("Limnodynastes dorsalis")),
                                 expression(italic("Limnodynastes peronii")),
                                 expression(italic("Litoria chloris")),
                                 expression(italic("Litoria xanthomera"))),
                     values = c(15,15,16,16,17,17)) +
  geom_smooth(aes(x = n_samples, y = value, 
                 color = as.factor(species),
                 linetype = as.factor(species)), 
              method = 'loess',
              formula = y ~ x,
              se = FALSE,
              size = 0.5, alpha = 1) +
  scale_linetype_manual(labels = c(expression(italic("Crinia insignifera")),
                                 expression(italic("Crinia parinsignifera")),
                                 expression(italic("Limnodynastes dorsalis")),
                                 expression(italic("Limnodynastes peronii")),
                                 expression(italic("Litoria chloris")),
                                 expression(italic("Litoria xanthomera"))),
                     values = c("solid", "solid", 
                                "longdash", "longdash", 
                                "dotted", "dotted")) +
  geom_vline(xintercept = 20, 
             color = "black", 
             linetype = "dashed", 
             size = 1.2) +
  theme_bw() +
  theme(text = element_text(color = "black", 
                            family = "sans",
                            size = 12),
        axis.text = element_text(color = "black", 
                            family = "sans",
                            size = 10),
        legend.text.align = 0,
        legend.position = "bottom",
        legend.title = element_blank()) +
  xlim(0,50) +
  facet_wrap(~ Variable, scale = "free") +
  ylab("Maximum Absolute Standard Deviation") +
  xlab("Number of Individuals in a Sample")

max_points
#ggsave(filename = "max_points.png",
 #      plot = max_points,
  #     path = "./exported_figures",
   #    device = "png",
    #   dpi = 2400,
     #  width = 7, height = 5)
```

