---
title: "Function to Get Call Variation Based on Area and Time"
authors: "Savannah Weaver, Corey Callaghan"
date: "January 01, 2020"
output: 
  github_document:
    toc: true
    toc_depth: 1
---

# Load Packages

```{r setup, echo=T, results='hide', message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if (!require("readr")) install.packages("readr")
library("readr")
if (!require("dplyr")) install.packages("dplyr")
library("dplyr")
if (!require("tidyr")) install.packages("tidyr")
library("tidyr")
if (!require("ggplot2")) install.packages("ggplot2")
library("ggplot2")
if (!require("sf")) install.packages("sf")
library("sf")
if (!require("lubridate")) install.packages("lubridate")
library("lubridate")
if (!require("lwgeom")) install.packages("lwgeom")
library("lwgeom")
if (!require("stringr")) install.packages("stringr")
library("stringr")
```


# Read In Data

```{r, echo=T, message=FALSE}
data <- read_csv("data_for_analysis/summary_data.csv")
```


# Make the Function

Calculate standard deviations of peak frequency and call duration, now as a function of time and geographic space. We run this for sample sizes 10, 20, 30, and 40, but only use n=20 later on. We measure the standard deviations, intra-annual difference in days, maximum geographic distance, and convex hull area (geographic area covered by the individuals in a given sample) for each sample.

The sd_variables_function calculates all these values:

```{r}
sd_variables_function <- function(draw, species_x){
  
  # do for each species
  sample_dat <- data %>%
    dplyr::filter(species==species_x) %>%
    ungroup() %>%
    sample_n(draw)
  
  # this part makes it so that if the genus is Litoria
  # then it takes the intra-annual difference across the season
  # this only really applies for litoria as the other two genera
  # have breeding seasons that are within a year
  genus <- word(species_x, 1)
  
  # get temporal difference of observations
  time_diff <- ifelse(genus == "Litoria", 
                      365 - max(yday(date(sample_dat$date)))+
                        min(yday(date(sample_dat$date))),
                      max(yday(date(sample_dat$date))) -
                        min(yday(date(sample_dat$date))))
  
  
  # convert FrogID points into sf format to use with polygons
  crs <- "+proj=longlat +ellps=GRS80 +no_defs"
  points_sf <- st_as_sf(sample_dat, 
                        coords = c("lng", "lat"), 
                        crs = crs, 
                        agr = "constant")
  
  # calculate polygon area for group of observations
  convex_hull <- st_convex_hull(st_union(points_sf))
  
  # calculate summary data and put all data into one dataframe
  summary_dat <- data.frame(sd_time=sd(sample_dat$ind_mean_time),
                            sd_freq=sd(sample_dat$ind_mean_freq)) %>%
    mutate(intra_annual_diff_days=time_diff) %>%
    mutate(convex_hull_area_km2=as.numeric(st_area(convex_hull))/1000) %>%
    mutate(max_distance_km=as.numeric(max(st_distance(points_sf)))/1000) %>%
    mutate(n_samples=draw) %>%
    mutate(centroid_lng=as.data.frame(st_coordinates(st_centroid(convex_hull)))$X) %>%
    mutate(centroid_lat=as.data.frame(st_coordinates(st_centroid(convex_hull)))$Y)
  
  # return dataframe of all calculated data
  return(summary_dat)
}
```

The boot_function_time_space repeats the sampling process 1000 times:

```{r}
boot_function_time_space <- function(draw, species_x){
  
  # rerun calculating variables for samples 1000 times
  # currently only 10 for testing purposes
  summary_dat <- do.call("rbind", 
                         purrr::rerun(1000, 
                                      sd_variables_function(draw, 
                                                            species_x)))
  
  # return the dataframe of all the reruns
  return(summary_dat)
}
```

The species_time_space_function applies the above functions to the 4 different sample sizes and to a given species:

```{r}
species_time_space_function <- function(species_name){
  
  # get data and do by species
  data_to_analyze <- data %>%
    dplyr::filter(species==species_name)
  
  # make a vector to run this function over
  ######NOTE: need to change to pick one sample size, probably 10 or 20
  vector_of_draws <- c(10, 20, 30, 40)
  
  # apply function to vector
  ##### We may not actually need this
  list_of_results <- lapply(vector_of_draws, 
                            boot_function_time_space, 
                            species_x=species_name)

  # put into dataframe
  df_of_results <- bind_rows(list_of_results) %>%
    mutate(species=species_name)
  
  # return dataframe with all calculated data
  return(df_of_results)
}
```

This code chunk runs the above functions for each species and saves the results:

```{r}
# do everything for each species
species_sd_correlation_list <- lapply(unique(data$species),
                                      function(x){species_time_space_function(x)})

# put all data for each species into one dataframe
species_sd_correlation_df <- bind_rows(species_sd_correlation_list) 

# export
saveRDS(species_sd_correlation_df, "data_for_analysis/standard_deviation_resampled_by_area_and_time.RDS")
```

The exported RDS file is analyzed in the "5_area_time_analysis.Rmd" file.