---
title: "Analysis of the Area Time Resampling Dataset"
authors: "Savannah Weaver, Corey Callaghan"
date: "01/01/2020"
output: 
  github_document:
    toc: true
    toc_depth: 1
---

# Load Packages

```{r setup, echo=T, results='hide', message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if (!require("tidyverse")) install.packages("tidyverse")
library("tidyverse")
if (!require("readr")) install.packages("readr")
library("readr")
if (!require("dplyr")) install.packages("dplyr")
library("dplyr")
if (!require("tidyr")) install.packages("tidyr")
library("tidyr")
if (!require("ggplot2")) install.packages("ggplot2")
library("ggplot2")
if (!require("lubridate")) install.packages("lubridate")
library("lubridate")
if (!require("lwgeom")) install.packages("lwgeom")
library("lwgeom")
if (!require("lme4")) install.packages("lme4")
library("lme4")
if (!require("lmerTest")) install.packages("lmerTest")
library("lmerTest")
if (!require("broom")) install.packages("broom")
library("broom")
if (!require("broom.mixed")) install.packages("broom.mixed")
library("broom.mixed")
if (!require("MASS")) install.packages("MASS")
library("MASS")
if (!require("purrr")) install.packages("purrr")
library("purrr")
if (!require("scales")) install.packages("scales")
library("scales")
if (!require("patchwork")) install.packages("patchwork")
library("patchwork")
if (!require("scales")) install.packages("scales")
library("scales")
library(extrafont)
loadfonts()
```


# Reading In Data

Get the datasets we made with bootstrapping, and check them for correct uploading.

```{r, echo=T, message=FALSE}
data <- read_rds("data_for_analysis/standard_deviation_resampled_by_area_and_time.RDS")
unique(data$n_samples)
```

We know we tried sample sizes of 10, 20, 30, and 40, so we can tell the data loaded in correctly.

Before we make plots, scale every samples variables (z-scores within species). We calculate mean and standard deviation for each species, then use those to calculate z-scores:

```{r}
scaled_area_days <- data %>%
  group_by(species, n_samples) %>%
  mutate(mean_sd_time = mean(sd_time)) %>%
  mutate(sd_sd_time = sd(sd_time)) %>%
  mutate(mean_sd_freq = mean(sd_freq)) %>%
  mutate(sd_sd_freq = sd(sd_freq)) %>%
  mutate(mean_diff_days = mean(intra_annual_diff_days)) %>%
  mutate(sd_diff_days = sd(intra_annual_diff_days)) %>%
  mutate(mean_area_km2 = mean(convex_hull_area_km2)) %>%
  mutate(sd_area_km2 = sd(convex_hull_area_km2)) %>%
  mutate(mean_max_km = mean(max_distance_km)) %>%
  mutate(sd_max_km = sd(max_distance_km)) %>%
  mutate(mean_lat = mean(centroid_lat)) %>%
  mutate(sd_lat = sd(centroid_lat)) %>%
  mutate(mean_lng = mean(centroid_lng)) %>%
  mutate(sd_lng = sd(centroid_lng)) %>%
  mutate(z_sd_time = (sd_time - mean_sd_time) / sd_sd_time) %>%
  mutate(z_sd_freq = (sd_freq - mean_sd_freq) / sd_sd_freq) %>%
  mutate(z_diff_days = (intra_annual_diff_days - mean_diff_days) / sd_diff_days) %>%
  mutate(z_area_km2 = (convex_hull_area_km2 - mean_area_km2) / sd_area_km2) %>%
  mutate(z_max_km = (max_distance_km - mean_max_km) / sd_max_km) %>%
  mutate(z_centroid_lat = (centroid_lat - mean_lat) / sd_lat) %>%
  mutate(z_centroid_lng = (centroid_lng - mean_lng) / sd_lng)
```

Also rename the most important variables to make recall easy:

```{r}
# prepare data
model_dat <- scaled_area_days %>%
  dplyr::filter(n_samples == 20) %>%
  rename(Frequency = "z_sd_freq") %>%
  rename(Time = "z_sd_time") %>%
  rename(Days = "z_diff_days") %>%
  rename(Area = "z_area_km2")
```

Check the response variable distributions before and after taking z-score:

```{r}
# sd freq
ggplot(model_dat, aes(x = sd_freq)) +
  geom_histogram(color = "black", fill = "darkseagreen", bins = 50) +
  theme_bw() +
  facet_wrap(~species, scales = "free") +
  xlab("Standard Deviation of Frequency")

# z sd freq
ggplot(model_dat, aes(x = Frequency)) +
  geom_histogram(color = "black", fill = "darkseagreen", bins = 50) +
  theme_bw() +
  facet_wrap(~species, scales = "free") +
  xlab("Z-Score Standard Deviation of Frequency")

# sd time
ggplot(model_dat, aes(x = sd_time))+
  geom_histogram(color = "black", fill = "darkseagreen", bins = 50)+
  theme_bw()+
  facet_wrap(~species, scales = "free")+
  xlab("Standard Deviation of Time")

# z sd time
ggplot(model_dat, aes(x = Time))+
  geom_histogram(color = "black", fill = "darkseagreen", bins = 50)+
  theme_bw()+
  facet_wrap(~species, scales = "free")+
  xlab("Z-Score Standard Deviation of Time")
```

Check the predictor variable distributions before and after taking z-score:

```{r}
# area
ggplot(model_dat, aes(x = convex_hull_area_km2)) +
  geom_histogram(color = "black", fill = "steelblue", bins = 50) +
  theme_bw() +
  facet_wrap(~species, scales = "free") +
  xlab(expression(paste("Convex Hull Area (km"^"2",")")))

# zscore area
ggplot(model_dat, aes(x = Area)) +
  geom_histogram(color = "black", fill = "steelblue", bins = 50) +
  theme_bw() +
  facet_wrap(~species, scales = "free") +
  xlab(expression(paste("Z-Score Convex Hull Area (km"^"2",")")))

# days
ggplot(model_dat, aes(x = intra_annual_diff_days)) +
  geom_histogram(color = "black", fill = "steelblue", bins = 50) +
  theme_bw() +
  facet_wrap(~species, scales = "free") +
  xlab("Intra-annual Difference (Days)")

# zscore days
ggplot(model_dat, aes(x = Days)) +
  geom_histogram(color = "black", fill = "steelblue", bins = 50) +
  theme_bw() +
  facet_wrap(~species, scales = "free") +
  xlab("Z-Score Intra-annual Difference (Days)")
```


# Plot sample SD as a function of **Area** and **Days**

In our plots, we only use the samples with n=20. 

```{r}
# call variables ~ area
scaled_area_days %>%
  dplyr::filter(n_samples == 20) %>%
  rename("Peak Frequency" = "sd_freq") %>%
  rename("Call Duration" = "sd_time") %>%
  gather(key = "Variable",
         value = "value",
         c("Peak Frequency", "Call Duration")) %>%
  ggplot(., aes(x = log10(convex_hull_area_km2),
                y = value,
                color = Variable)) +
  geom_smooth(method = "lm",
              formula = y ~ x) +
  scale_colour_manual(values = c(rep(c(rgb(168/255,221/255,181/255),
                                       rgb(67/255,162/255,202/255)),3))) +
  facet_wrap(species ~ Variable, scales = "free") +
  ylab("Standard Deviation") +
  xlab(expression(paste("log10 Geographic Area (km"^"2",")"))) +
  theme_bw() +
  scale_x_continuous(breaks = scales::pretty_breaks(3)) +
  scale_y_continuous(breaks = scales::pretty_breaks(3)) +
  theme(text = element_text(color = "black", 
                            family = "sans",
                            size = 12),
        axis.text = element_text(color = "black", 
                            family = "sans",
                            size = 10),
        strip.background = element_rect(color = "white", 
                                        fill = "white", 
                                        size = 1, 
                                        linetype = "solid"),
        strip.text = element_text(color = "black", 
                            family = "sans",
                            size = 10),
        legend.position = "none") -> call_var_by_area

call_var_by_area
# export plot
#ggsave(filename =  "geographic_area_lines.png",
 #      plot = call_var_by_area,
  #     path = "exported_figures",
   #    device = "png",
    #   dpi = 1200,
     #  width = 7, height = 6)



# call variables ~ days
scaled_area_days %>%
  dplyr::filter(n_samples == 20) %>%
  rename("Peak Frequency" = "sd_freq") %>%
  rename("Call Duration" = "sd_time") %>%
  gather(key = "Variable",
         value = "value",
         c("Peak Frequency", "Call Duration")) %>%
  ggplot(., aes(x = intra_annual_diff_days,
                y = value,
                color = Variable)) +
  geom_smooth(method = "lm",
              formula = y ~ x) +
  scale_colour_manual(values = c(rep(c(rgb(168/255,221/255,181/255),
                                       rgb(67/255,162/255,202/255)),3))) +
  facet_wrap(species ~ Variable, scales = "free") +
  ylab("Standard Deviation") +
  xlab("Intra-annual Time Difference (Days)") +
  theme_bw() +
  scale_x_continuous(breaks = scales::pretty_breaks(4)) +
  scale_y_continuous(breaks = scales::pretty_breaks(3)) +
  theme(text = element_text(color = "black", 
                            family = "sans",
                            size = 12),
        axis.text = element_text(color = "black", 
                            family = "sans",
                            size = 10),
        strip.background = element_rect(color = "white", 
                                        fill = "white", 
                                        size = 1, 
                                        linetype = "solid"),
        strip.text = element_text(color = "black", 
                            family = "sans",
                            size = 10),
        legend.position = "none") -> call_var_by_days

call_var_by_days
# export plot
#ggsave(filename =  "intra_annual_days_lines.png",
 #      plot = call_var_by_days,
  #     path = "exported_figures",
   #    device = "png",
    #   dpi = 1200,
     #  width = 7, height = 6)
```


# Model Relationships

We do a global mixed model for each combination of: 
*predictor variables*: area, days
*response variables*: z-score of sample peak frequency standard deviation,
z-score of sample call time duration standard deviation. Species is included in the model.

```{r}
# time ~ area
mixed_mod_time_area <- lmer(Time ~ Area + (1|species), 
                            data = model_dat)
AT <- broom.mixed::tidy(mixed_mod_time_area)
AT$predictor <- "Area"
AT$response <- "Time"

# time ~ days
mixed_mod_time_days <- lmer(Time ~ Days + (1|species), 
                            data = model_dat)
DT <- broom.mixed::tidy(mixed_mod_time_days)
DT$predictor <- "Days"
DT$response <- "Time"

# freq ~ area
mixed_mod_freq_area <- lmer(Frequency ~ Area + (1|species), 
                            data = model_dat)
AF <- broom.mixed::tidy(mixed_mod_freq_area)
AF$predictor <- "Area"
AF$response <- "Frequency"

# freq ~ days
mixed_mod_freq_days <- lmer(Frequency ~ Days + (1|species), 
                            data = model_dat)
DF <- broom.mixed::tidy(mixed_mod_freq_days)
DF$predictor <- "Days"
DF$response <- "Frequency"

# put all into one dataframe
all_mixed_mods <- AT %>%
  bind_rows(.,DT,AF,DF) %>%
  dplyr::filter(effect == "fixed")

# save mixed model data
write_csv(all_mixed_mods, "exported_tables/all_mixed_mods.csv")
```

Model for **each species separately**:

```{r, message=FALSE}
# time ~ area
time_area_models <- model_dat %>%
  nest(-species) %>%
  mutate(
    fit = map(data, ~ lm(Time ~ Area, data = .x)),
    tidied = map(fit, tidy),
    glanced = map(fit, glance),
    confint = map(fit, confint_tidy)
  )

time_area_summary <- time_area_models %>%
  unnest(tidied) %>%
  dplyr::select(-data, -fit, -glanced, -confint) %>%
  left_join(., time_area_models %>%
              unnest(glanced) %>%
              dplyr::select(-data, -fit, -tidied, -confint), 
            by = "species") %>%
  left_join(., time_area_models %>%
              unnest(confint) %>%
              dplyr::select(-data, -fit, -tidied, -glanced) %>%
              group_by(species) %>%
              slice(2), by = "species") %>%
  mutate(model = "time_vs_area")

# time ~ days
time_days_models <- model_dat %>%
  nest(-species) %>%
  mutate(
    fit = map(data, ~ lm(Time ~ Days, data = .x)),
    tidied = map(fit, tidy),
    glanced = map(fit, glance),
    confint = map(fit, confint_tidy)
  )

time_days_summary <- time_days_models %>%
  unnest(tidied) %>%
  dplyr::select(-data, -fit, -glanced, -confint) %>%
  left_join(., time_days_models %>%
              unnest(glanced) %>%
              dplyr::select(-data, -fit, -tidied, -confint), 
            by = "species") %>%
  left_join(., time_days_models %>%
              unnest(confint) %>%
              dplyr::select(-data, -fit, -tidied, -glanced) %>%
              group_by(species) %>%
              slice(2), by = "species") %>%
  mutate(model = "time_vs_days")

# freq ~ area
freq_area_models <- model_dat %>%
  nest(-species) %>%
  mutate(
    fit = map(data, ~ lm(Frequency ~ Area, data = .x)),
    tidied = map(fit, tidy),
    glanced = map(fit, glance),
    confint = map(fit, confint_tidy)
  )

freq_area_summary <- freq_area_models %>%
  unnest(tidied) %>%
  dplyr::select(-data, -fit, -glanced, -confint) %>%
  left_join(., freq_area_models %>%
              unnest(glanced) %>%
              dplyr::select(-data, -fit, -tidied, -confint), 
            by = "species") %>%
  left_join(., freq_area_models %>%
              unnest(confint) %>%
              dplyr::select(-data, -fit, -tidied, -glanced) %>%
              group_by(species) %>%
              slice(2), by = "species") %>%
  mutate(model = "freq_vs_area")

# freq ~ days
freq_days_models <- model_dat %>%
  nest(-species) %>%
  mutate(
    fit = map(data, ~ lm(Frequency ~ Days, data = .x)),
    tidied = map(fit, tidy),
    glanced = map(fit, glance),
    confint = map(fit, confint_tidy)
  )

freq_days_summary <- freq_days_models %>%
  unnest(tidied) %>%
  dplyr::select(-data, -fit, -glanced, -confint) %>%
  left_join(., freq_days_models %>%
              unnest(glanced) %>%
              dplyr::select(-data, -fit, -tidied, -confint), 
            by = "species") %>%
  left_join(., freq_days_models %>%
              unnest(confint) %>%
              dplyr::select(-data, -fit, -tidied, - glanced) %>%
              group_by(species) %>%
              slice(2), by = "species") %>%
  mutate(model = "freq_vs_days")

# put the model results together and add range data
all_model_summaries <- time_area_summary %>%
  bind_rows(time_days_summary) %>%
  bind_rows(freq_area_summary) %>%
  bind_rows(freq_days_summary) %>%
  dplyr::filter(term != "(Intercept)") %>%
  dplyr::select(-statistic.y, -p.value.y) %>%
  rename(statistic = statistic.x) %>%
  rename(p.value = p.value.x) %>%
  left_join(., read_csv("data_for_analysis/species_ranges.csv") %>%
              rename(species=Species))

# save
#write_csv(all_model_summaries, "exported_tables/each_species_model_summaries.csv")
```

Plot models by species:

```{r}
all_model_summaries %>%
  mutate(term = case_when(term == "Area" ~ "Geographic Area",
                          term == "Days" ~ "Intra-annual Time Difference")) %>%
  mutate(model = case_when(model == "freq_vs_area" ~ "Peak Frequency",
                         model == "freq_vs_days" ~ "Peak Frequency",
                         model == "time_vs_area" ~ "Call Duration",
                         model == "time_vs_days" ~ "Call Duration"
  )) %>%
  ggplot(., aes(x = species, 
                y = estimate, 
                color = model)) +
  geom_hline(yintercept = 0, 
             color = "black", 
             linetype = "dashed") +
  geom_point(size = 2,
             position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(ymin = conf.low, 
                    ymax = conf.high, 
                    color = model), 
                width = 0.4, 
                position = position_dodge(width = 0.5)) +
  coord_flip() +
  theme_bw() +
  facet_wrap(~ term, scales = "free_x") +
  xlab("") +
  ylab("Parameter Estimate") +
  theme(text = element_text(color = "black", 
                            family = "sans",
                            size = 12),
        axis.text = element_text(color = "black", 
                            family = "sans",
                            size = 10),
        axis.text.y = element_text(face = "italic"),
        legend.title = element_blank()) +
  scale_colour_manual(values = c(rep(c(rgb(168/255,221/255,181/255),
                                       rgb(67/255,162/255,202/255)),3))) +
  labs(color = "Response: ") +
  theme(legend.position = "bottom") -> model_est_points

model_est_points
# export plot
#ggsave(filename = "model_est_points.png",
 #      plot = model_est_points,
  #     path = "./exported_figures",
   #    device = "png",
    #   dpi = 2400,
     #  width = 7, height = 4)
```


# Correlation and Species Range

Does range size mean anything for these results? Does the strength of the above tested correlations vary based on a species range size?

```{r}
# (call duration ~ days) ~ range
time_days_range <- all_model_summaries %>% 
  dplyr::filter(model == "time_vs_days")
time_days_range_model <- glm(data = time_days_range,
                             estimate ~ range_size_km2)
summary(time_days_range_model)

# (call duration ~ area) ~ range
time_area_range <- all_model_summaries %>% 
  dplyr::filter(model == "time_vs_area")
time_area_range_model <- glm(data = time_area_range,
                             estimate ~ range_size_km2)
summary(time_area_range_model)

# (frequency ~ days) ~ range
freq_days_range <- all_model_summaries %>% 
  dplyr::filter(model == "freq_vs_days")
freq_days_range_model <- glm(data = freq_days_range,
                             estimate ~ range_size_km2)
summary(freq_days_range_model)

# (frequency ~ area) ~ range
freq_area_range <- all_model_summaries %>% 
  dplyr::filter(model == "freq_vs_area")
freq_area_range_model <- glm(data = freq_area_range,
                             estimate ~ range_size_km2)
summary(freq_area_range_model)
```


# Correlation and Species Body Size

Does body size mean anything for these results? Does vocal variability depend on maximum potential body size? 

Add body size data:

```{r}
model_sum_svl <- all_model_summaries %>%
  left_join(., read_csv("data_for_analysis/body_sizes.csv"),
                        by = "species") %>%
  rename(max_male_SVL_mm = male_body_size_max_mm)
```

Model:

```{r}
# (call duration ~ days) ~ SVL
time_days_body <- model_sum_svl %>% 
  dplyr::filter(model == "time_vs_days")
time_days_body_model <- glm(data = time_days_body,
                             estimate ~ max_male_SVL_mm)
summary(time_days_body_model)

# (call duration ~ area) ~ SVL
time_area_body <- model_sum_svl %>% 
  dplyr::filter(model == "time_vs_area")
time_area_body_model <- glm(data = time_area_body,
                             estimate ~ max_male_SVL_mm)
summary(time_area_body_model)

# (frequency ~ days) ~ SVL
freq_days_body <- model_sum_svl %>% 
  dplyr::filter(model == "freq_vs_days")
freq_days_body_model <- glm(data = freq_days_body,
                             estimate ~ max_male_SVL_mm)
summary(freq_days_body_model)

# (frequency ~ area) ~ SVL
freq_area_body <- model_sum_svl %>% 
  dplyr::filter(model == "freq_vs_area")
freq_area_body_model <- glm(data = freq_area_body,
                             estimate ~ max_male_SVL_mm)
summary(freq_area_body_model)
```

